\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-11-15 smurf_paper.tex
Add vertical limits to caption for Fig.9
didn't quite do the 3-way merge correctly...

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting nominally of 5120 bolometers (with a useful yield of
  69\%) in each of two simultaneous imaging bands centred over 450 and
  850\,\micron. The camera is operated by scanning across the sky and
  recording data at a rate of 200\,Hz. As the largest of a new
  generation of multiplexed kilopixel bolometer cameras operating in
  the (sub)millimetre, \scuba\ data analysis represents a significant
  challenge.  We describe the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of science
  products for the JCMT Science Archive at the Canadian Astronomy Data
  Centre. Three detailed case studies are used to: (i) explore
  convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $10,240$ nominal detectors \citep[of which $\sim$69\%
work and are typically useful,][]{holland2012}, \scuba\ is presently
the largest of a new generation of multiplexed kilopixel
(sub)millimetre bolometer cameras, which also includes the cameras for
the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
experiments for studying anisotropies in the Cosmic Microwave
Backround (CMB) using similar technology.

@@

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is scanned at a
range of position angles, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$), such that significant drift in the
bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles. Such methods are now used by virtually all existing
ground-based submillimetre camerass
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
preference to ``chopping'' methods (where the secondary is moved
quickly to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensitive to modest angular scales. We note that under some
circumstances cross-linking may not be essential. For example, SPT,
which was designed to measure CMB anisotropies, scans entirely in
azimuth \citep{schaffer2011}. While this strategy limits its ability
to recover large angular scales transverse to the scan direction, the
anisotropic filtering in its maps can be accounted for during analysis
to achieve SPTs focussed objectives \citep[in contrast, the similar
ACT experiment uses cross-linking to improve its reponse to
large-angular scales,][]{das2011}. This approach is less practical as
a solution for \scuba, however, which must serve a broader range of
scientific interests.

@@

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the experimental CMB literature
include \citet{tegmark1997} and \citet{stompor2002}, and an
application to data from SCUBA is described in \citet{borys2004}.  The
downside to such methods is that they can be both computationally
expensive and have large memory requirements. While for some
experimental designs fast iterative methods for the inversion do exist
without requiring excessive amounts of memory such as the method of
\citet{wright1996} \citep[which was later implemented for SCUBA
by][]{johnstone2000}, for the more general map-making problem,
involving many detectors, matters are significantly complicated both
by the need to measure the cross spectra for all unique pairs of
bolometers, as well as performing the inversion itself. A
maximum-likelihood method was successfully used to make maps from ACT
data using approximately the same number of detectors as a single
\scuba\ subarray at a single wavelength. However, the calculation is
extremely resource intensive and requires a large cluster in order to
run; a single map can take 10\,CPU\,years \citep{fowler2010}. Perhaps
the most promising maximum-likelihood method that may one day be
applied to \scuba\ data is ``SANEPIC'', which produced maps from
Balloon-borne Large-Aperture Submillimeter Telescope data, involving
hundreds of bolometers, while correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

@@

Iterative approaches to map-making have a long history in the submillimetre. For example, the iterative pipeline recipe for fitting and
removing baseline drifts in SCUBA scan-map data as the astronomical
image estimate improved \citep{1999ASPC..172..171J} is similar to the
algorithm described in this paper. The closest modern relatives are
the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
Bolometer array data Analysis software (BoA) for the Large APEX
BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
developed for the Bolocam Galactic Plane Survey \citep{aguirre2011}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales comensurate with the observing
times (or significantly faster), on single high-end desktop
computers. For reference, all of the data analysis in this paper was
performed on a machine with 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline offering feedback to
observers at the telescope. The pipeline is also used to generate
near-science grade products for the JCMT Science Archive
\citep{2011ASPC..442..203E} hosted by the Canadian Astronomy Data
Centre (CADC). Details of the pipeline design are given in
\citet{gibb2005} and \citet{2008AN....329..295C}.

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry.
Temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields, which are detected and amplified
using chains of superconducting quantum interference devices (SQUIDs),
before the larger output currents are digitised. Each detector has its
own SQUID for the first-stage of the amplification, but the remaining
stages occur within a common chain of SQUIDs for each of the 32
columns. All 40 rows are read out in sequence at a row-visit frequency
of about 12\,kHz. Such a high sample rate is unnecessary to produce
maps, so the data are re-sampled to approximately 200\,Hz before
writing to disk. This rate provides a sample every 3.0\,arcsec, or
approximately one third of the 450\,\micron\ diffraction-limited
full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
adequately sampling a Gaussian point spread function (PSF) -- at the
maximum scanning speed of 600\,arcsec\,s$^{-1}$. There is an
additional 41st row of SQUID readouts that are not connected to
TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
common to each column's amplifier chain. The relationship between the
output digitised current, $I$, and the input power, $P$, or $dI/dP$ is
established using flatfield observations immediately prior to science
observations, in which the output signal is measured throughout a ramp
of the pixel heaters (which provide a known input power); see
Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012} for more details. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the JCMT Water Vapour
Monitor to track line-of-sight opacity variations, see Section~3
in][]{dempsey2012}, and the application of a flux conversion factor
(FCF) which is established from regular measurements of calibrators
such as Uranus \citep[Section~5 in][]{dempsey2012}.

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra include a roll-off in their spectra at frequencies approaching
Nyquist, which is due to the anti-aliasing filter that forms part of
the 200\,Hz re-sampling stage. There are also line features which are
thought to be produced at frequencies far above the final 200\,Hz
sample rate, and are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area on several
different time scales, and at different scan angles); and (ii) minimal
accelerations to reduce turn-around overheads (which could be quite
large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
larger than the array footprint a rectangular ``PONG'' pattern is
used, in which the boresight travels in approximately straight lines
and ``bounces'' off the edges at 45\,degree angles until the area is
uniformly filled in. It is also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking. For smaller areas (of order the
array footprint, or point sources), in which the PONG turn-around
overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
used.  Here the telescope moves in a circle, whose centre also slowly
traces out a small circle. For a more complete description of the
\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

@@

The low-frequency signal component of the S2SRO bolometer output is
also highly correlated amongst bolometers in the same subarray. We
have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
time-series of all the working bolometers. We then fit the amplitude
of $\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are quite flat, although still with
noticeable variations. The white noise is also apparent, and larger at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (however, note that variations are still seen in the mixing
chamber signal; such variations do not necessarily reflect changes in
the focal plane temperature). There is minimal correlated
low-frequency noise in these 450\,\micron\ data (note that there is
almost no difference once the common-mode has been subtracted). The
850\,\micron\ channel, however, exhibits a more significant signal
that is obviously correlated with the telescope motion. It is also
correlated amongst many of the detectors, and common-mode removal
corrects it to some extent. Part of the reason that this is seen here,
and not in the S2SRO data set shown, is that the amplitude of the scan
pattern is larger. Most of this ``scan-synchronous'' noise is
attributed to magnetic field pickup, as described in
Section~\ref{sec:magpickup}.

@@

\item Apodisation: a number of samples at the start and end of each
bolometer time-series are multiplied by a cosine function in order to
roll the data values off smoothly to zero. The default number of
samples modified at each end of the time-series is given by half the
ratio of the sampling frequency to the lowest retained frequency. In
addition, each end of the time-series is padded with double this
number of zeros. This method is illustrated in Fig.~\ref{fig:pad2}. It
is not used by default as it reduces the amount of data available for
the map, and can significantly hinder very short observations (e.g.,
of calibrators when focussing the telescope).

@@

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com}
\caption{A selection of typical bolometer time-series after
flatfielding and removal of a constant baseline. Each sub-plot shows 
the power in pW from a single bolometer over a single 30 second period.  
It can be seen that most bolometers exhibit a common time variation overlaid with other
features. Empty squares indicate the locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

As described in Section~\ref{sec:magpickup}, wide scans can produce
significant magnetic field pickup that tracks the azimuthal motion of
the telescope. The \model{DKS} model uses the dark squid signals for
each column as a template that is simply scaled (gain and offset) to
each bolometer time-series before removal. Unfortunately the dark
squids do not work for every column (several in each subarray),
meaning that those columns are usually discarded when producing
maps. One possible remedy is to identify dead bolometers (with
otherwise working TES readouts) or intentionally disconnect working
bolometers, to create replacement dark readouts in those columns. This
solution has not been pursued due to the limited success at removing
pickup using the presently working dark squids. Another alternative
model is \model{TMP} in which the azimuth of the telescope itself is
used as the template. For both \model{DKS} and \model{TMP}, there is
some success at visibly decreasing the low-frequency scan-synchronous
noise in certain data sets, although often it does not (and there are
also examples in which the noise is increased). In the case of
\model{DKS}, there may simply be components of the signal seen by
working detectors that are not detecting in the dark
squids. Furthermore, in the case of \model{TMP}, the relationship
between the projection of the Earth's magnetic field on the instrument
and the magnitude of the pickup may not be linearly correlated as we
have assumed (verified in some cases by a comparison of the azimuthal
motion with working dark squid signals). In no case was the use of
these models able to reduce the $1/f$ knee substantially,
necessitating the continued use of \model{FLT} to remove the remaining
low-frequency noise. Neither \model{DKS} nor \model{TMP} are typically
used.

@@

The first, simplest reduction of the data uses only the \model{COM}
model to estimate and remove the common-mode signal in order to
suppress low-frequency noise in the data. After \model{COM}, the
extinction correction is applied (\model{EXT}), and an initial map is
estimated using equal weighting for all of the detectors. This
estimate of \model{AST} is then removed from the data, and the noise
is measured in the residuals to estimate weights for the subsequent
and final iteration. The resulting map after these two iterations is
shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
clearly large ($\sim$250 as estimated by SMURF), enabling us to see
the faint sidelobes (circle and cross pattern), the map also has
obvious circular streaks and other large-scale ripples. The circular
streaks are due to the fact that \model{COM} does not account for all
of the low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore
each bolometer leaves a trace of the circular scan pattern in the map,
as each of their baselines slowly drift independently. A significant
contribution to the larger-scale ripples in the map, however, can be
made by degeneracies in the map solution, as we discuss below.

@@

One improvement to remove residual low-frequency noise after
common-mode removal in this Uranus image, is to simply apply a
high-pass filter to the data. A third reduction of the data uses the
default map-making parameters, as described in
Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
accomplish this task immediately prior to map estimation. We set the
filter edge based on an angular scale of 200\,arcsec, which, given the
scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
0.78\,Hz. Now using the automatic map-based convergence test
(Section~\ref{sec:converge}), the resulting map converges after 10
iterations (Fig.~\ref{fig:pointmaps}c). Both the circular streaks and
the large-scale gradient are now removed, but they have been replaced
by an obvious, circularly-symmetric ringing pattern about Uranus. The
reason for this ringing is that the hard-edged high-pass filter in
frequency space is equivalent to a $\sinc$ function-like response in
map space. Since the scan pattern is fairly isotropic (scans at all
position angles), and there is a bright point-like source at the
centre, the result is an azimuthally-symmetric $\sinc$ function-like
pattern in the map.

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers (with a useful yield of
  69\%) in each of two simultaneous imaging bands centred over 450 and
  850\,\micron. The camera is operated by scanning across the sky and
  recording data at a rate of 200\,Hz. As the largest of a new
  generation of multiplexed kilopixel bolometer cameras operating in
  the (sub)millimetre, \scuba\ data analysis represents a significant
  challenge.  We describe the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of science
  products for the JCMT Science Archive at the Canadian Astronomy Data
  Centre. Three detailed case studies are used to: (i) explore
  convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $10,240$ nominal detectors \citep[of which $\sim$69\%
work and are typically useful,][]{holland2012}, \scuba\ is presently
the largest of a new generation of multiplexed kilopixel
(sub)millimetre bolometer cameras, which also includes the cameras for
the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
experiments for studying anisotropies in the Cosmic Microwave
Backround (CMB) using similar technology, the latter of which uses the
same time-domain multiplexed readout electronics as \scuba\
\citep{battistelli2008}.

@@

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is scanned at a
range of position angles, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$), such that significant drift in the
bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles. Such methods are now used by virtually all existing
ground-based submillimetre camerass
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
preference to ``chopping'' methods (where the secondary is moved
quickly to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensitive to modest angular scales. We note that under some
circumstances cross-linking may not be essential. For example, SPT,
which was designed to measure CMB anisotropies, scans entirely in
azimuth \citep{schaffer2011}. While this strategy limits its ability
to recover large angular scales transverse to the scan direction, the
anisotropic filtering in its maps can be accounted for during analysis
to achieve SPTs focussed objectives \citep[in contrast, the similar
ACT experiment uses cross-linking to improve its reponse to
large-angular scales,][]{das2011}. This approach is less practical as
a general solution for \scuba, however, which must serve a broader
range of scientific interests.

@@

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the experimental CMB literature
include \citet{tegmark1997} and \citet{stompor2002}, and an
application to data from SCUBA is described in \citet{borys2004}.  The
downside to such methods is that they can be both computationally
expensive and have large memory requirements. While for some
experimental designs fast iterative methods for the inversion do exist
without requiring excessive amounts of memory, such as that of
\citet{wright1996} \citep[which was later implemented for SCUBA
by][]{johnstone2000}, for the more general map-making problem,
involving many detectors, matters are significantly complicated both
by the need to measure the cross spectra for all unique pairs of
bolometers, as well as performing the inversion itself. A
maximum-likelihood method was successfully used to make maps from ACT
data using approximately the same number of detectors as a single
\scuba\ subarray at a single wavelength. However, the calculation is
extremely resource intensive and requires a large cluster in order to
run; a single map can take 10\,CPU\,years \citep{fowler2010}. Perhaps
the most promising maximum-likelihood method that may one day be
applied to \scuba\ data is ``SANEPIC'', which produced maps from
Balloon-borne Large-Aperture Submillimeter Telescope data, involving
hundreds of bolometers, while correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

@@

Iterative approaches to map-making have a long history in the
submillimetre, such as the pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The closest modern
relatives are the Comprehensive Reduction Utility
\citep[CRUSH,][]{kovacs2008} for the Submillimeter High-Angular
Resolution Camera 2 (SHARC-2), the Bolometer array data Analysis
software (BoA) for the Large APEX BOlometer CAmera
\citep[LABOCA,][]{siringo2009}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2011}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales comensurate with the observing
times (or significantly faster), on single high-end desktop
computers. For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline offering feedback to
observers at the telescope. The pipeline is also used to generate
near-science grade products for the JCMT Science Archive
\citep{2011ASPC..442..203E} hosted by the Canadian Astronomy Data
Centre (CADC). Details of the pipeline design are given in
\citet{gibb2005} and \citet{2008AN....329..295C}.

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers (with a response
time constant of $\tau \sim 1$\,ms) coupled to superconducting
transition-edge sensors (TESs) for thermometry. Temperature variations
in the TESs produce changing currents, and therefore varying magnetic
fields, which are detected and amplified using chains of
superconducting quantum interference devices (SQUIDs), before the
larger output currents are digitised. Each detector has its own SQUID
for the first-stage of the amplification, but the remaining stages
occur within a common chain of SQUIDs for each of the 32 columns. All
40 rows are read out in sequence at a row-visit frequency of about
12\,kHz. Such a high sample rate is unnecessary to produce maps, so
the data are re-sampled to approximately 200\,Hz before writing to
disk. This rate provides a sample every 3.0\,arcsec, or approximately
one third of the 450\,\micron\ diffraction-limited full-width at
half-maximum (FWHM) -- a typical rule-of-thumb for adequately sampling
a Gaussian point spread function (PSF) -- at the maximum scanning
speed of 600\,arcsec\,s$^{-1}$. There is an additional 41st row of
SQUID readouts that are not connected to TESs. These ``dark SQUIDs''
track non-thermal noise sources that are common to each column's
amplifier chain. The relationship between the output digitised
current, $I$, and the input power, $P$, or $dI/dP$ is established
using flatfield observations immediately prior to science
observations, in which the output signal is measured throughout a ramp
of the pixel heaters (which provide a known input power); see
Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012} for more details. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the JCMT Water Vapour
Monitor to track line-of-sight opacity variations, see Section~3
in][]{dempsey2012}, and the application of a flux conversion factor
(FCF) which is established from regular measurements of calibrators
such as Uranus \citep[Section~5 in][]{dempsey2012}.

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim7$\,ms (which must be added to the 1\,ms thermal
time constant of the bolometers mentioned above to obtain the total
lag). There are also line features which are thought to be produced at
frequencies far above the final 200\,Hz sample rate, and are aliased
to lower frequencies during the multiplexed readout stage before they
can be removed by the anti-aliasing filter.

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area at different
scan angles); and (ii) minimal accelerations to reduce turn-around
overheads (which could be quite large given the 600\,arcsec\,s$^{-1}$
maximum scan speed). For areas larger than the array footprint a
rectangular ``PONG'' pattern is used, in which the boresight travels
in approximately straight lines and ``bounces'' off the edges at
45\,degree angles until the area is uniformly filled in. It is also
usually combined with a rotation through a number of fixed position
angles to create a ``rotating PONG'' with even better
cross-linking. For smaller areas (of order the array footprint, or
point sources), in which the PONG turn-around overheads would be
large, a ``constant velocity daisy'' (CV Daisy) is used.  Here the
telescope moves in a circle, whose centre also slowly traces out a
small circle. For a more complete description of the \scuba\ observing
modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

@@

The low-frequency signal component of the S2SRO bolometer output is
also highly correlated amongst bolometers in the same subarray. We
have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
time-series of all the working bolometers. We then fit the amplitude
of $\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are nearly white, although still with
noticeable long-timescale variations. The white noise is greater at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (however, note that variations are still seen in the mixing
chamber signal; such variations do not necessarily reflect changes in
the focal plane temperature). There is minimal correlated
low-frequency noise in these 450\,\micron\ data (note that there is
almost no difference once the common-mode has been subtracted). The
850\,\micron\ channel, however, exhibits a more significant signal
that is obviously correlated with the telescope motion. It is also
correlated amongst many of the detectors, and common-mode removal
corrects it to a large extent. Part of the reason that this is seen
here, and not in the S2SRO data set shown, is that the amplitude of
the scan pattern is larger. Most of this ``scan-synchronous'' noise is
attributed to magnetic field pickup, as described in
Section~\ref{sec:magpickup}.

@@

\item Apodisation: a number of samples at the start and end of each
bolometer time-series are multiplied by a cosine function in order to
roll the data values off smoothly to zero. The default number of
samples modified at each end of the time-series is given by half the
ratio of the sampling frequency to the lowest retained frequency, the
edge frequency of the high-pass filter (Section~\ref{sec:flt}). In
addition, each end of the time-series is padded with double this
number of zeros. This method is illustrated in Fig.~\ref{fig:pad2}. It
is not used by default as it reduces the amount of data available for
the map, and can significantly hinder very short observations (e.g.,
of calibrators when focussing the telescope).

@@

In principle there are two other effects that could be included in
this model. First, signals detected by the bolometers (inside the
brackets of Eq.~\ref{eq:model}) should be convolved with the thermal
response function, $\exp(-t/\tau)$, where $\tau=1$\,ms. Second, the
final signals should be convolved with the response of the
anti-aliasing filter. In practice, the combined effect of these two
filters is to introduce a lag of $\sim$7\,ms, and only minimal
attentuation of a point source (at most 3\% for the smaller
450\,\micron\ PSF at the maximum scan speed of 600\,arcsec). For all
of the maps described in this paper, their impact is therefore
negligible. The current version of SMURF (part of the Starlink
\textsc{kapuahi} release) does not attempt to de-convolve these
effects, although the next version will likely include an optional
pre-processing step to perform this task.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com}
\caption{A selection of typical bolometer time-series after
flatfielding and removal of a constant baseline. Each sub-plot shows
the power from a single bolometer over a 30 second period.
The lower and upper vertical limits of each plot are -0.02 pW and +0.02
pW respectively. It can be seen that most bolometers exhibit a common
time variation overlaid with other features. Empty squares indicate the
locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

As described in Section~\ref{sec:magpickup}, wide scans can produce
significant magnetic field pickup that tracks the azimuthal motion of
the telescope. The \model{DKS} model uses the dark squid signals for
each column as a template that is simply scaled (gain and offset) to
each bolometer time-series before removal. Unfortunately the dark
squids do not work for every column (several cases in each subarray),
meaning that those columns are usually discarded when producing
maps. One possible remedy is to identify dead bolometers (with
otherwise working TES readouts) or intentionally disconnect working
bolometers, to create replacement dark readouts in those columns. This
solution has not been pursued due to the limited success at removing
pickup using the presently working dark squids. Another alternative
model is \model{TMP} in which the azimuth of the telescope itself is
used as the template. For both \model{DKS} and \model{TMP}, there is
some success at visibly decreasing the low-frequency scan-synchronous
noise in certain data sets, although often it does not (and there are
also examples in which the noise is increased). In the case of
\model{DKS}, there may simply be components of the signal seen by
working detectors that are not apparent in the dark
squids. Furthermore, in the case of \model{TMP}, the relationship
between the projection of the Earth's magnetic field on the instrument
and the magnitude of the pickup may not be linearly correlated as we
have assumed (verified in some cases by a comparison of the azimuthal
motion with working dark squid signals). In no case was the use of
these models able to reduce the $1/f$ knee substantially,
necessitating the continued use of \model{FLT} to remove the remaining
low-frequency noise. Neither \model{DKS} nor \model{TMP} are typically
used.

@@

The first, simplest reduction of the data uses only the \model{COM}
model to estimate and remove the common-mode signal in order to
suppress low-frequency noise in the data. After \model{COM}, the
extinction correction is applied (\model{EXT}), and an initial map is
estimated using equal weighting for all of the detectors. This
estimate of \model{AST} is then removed from the data, and the noise
is measured in the residuals to estimate weights for the subsequent
and final iteration. The resulting map after these two iterations is
shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
clearly large ($\sim$250), enabling us to see the faint sidelobes
(circle and cross pattern), the map also has obvious circular streaks
and other large-scale ripples. The circular streaks are due to the
fact that \model{COM} does not account for all of the low-frequency
(i.e., uncorrelated) noise (see Fig.~\ref{fig:pspec}), and therefore
each bolometer leaves a trace of the circular scan pattern in the map,
as their baselines slowly drift independently. A significant
contribution to the larger-scale ripples in the map, however, can be
made by degeneracies in the map solution, as we discuss below.

@@

Attacking the problem of streaks, a simple method of removing residual
(un-correlated) sources of low-frequency noise is to apply a high-pass
filter after the common-mode removal. A third reduction of the data
uses the default map-making parameters, as described in
Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
accomplish this task immediately prior to map estimation. We set the
filter edge based on an angular scale of 200\,arcsec, which, given the
scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
0.78\,Hz (note in Fig.~\ref{fig:pspec} that this frequency is slightly
above the $1/f$ knee at 450\,\micron\ after common-mode removal). Now
using the automatic map-based convergence test
(Section~\ref{sec:converge}), the solution converges after 10
iterations (Fig.~\ref{fig:pointmaps}c). Both the circular streaks and
the large-scale gradient are now removed, but they have been replaced
by an obvious, circularly-symmetric ringing pattern about Uranus. The
reason for this ringing is that the hard-edged high-pass filter in
frequency space is equivalent to a $\sinc$ function-like response in
map space. Since the scan pattern is fairly isotropic (scans at all
position angles), and there is a bright point-like source at the
centre, the result is an azimuthally-symmetric $\sinc$ function-like
pattern in the map. The gradient resulting from the degeneracy between
the common-mode and map has been removed because the filter scale of
200\,arcsec (3.3\,arcmin) is smaller than the array field-of-view
($\sim$5\,arcmin). If it were greater, degeneracies would again begin
to appear on those larger scales.

This example illustrates the need for constraints in the map solution
in many situations. For calibrators (and other previously known
bright, compact sources), a good, simple prior is to constrain the map
to a value of zero away from the known locations of emission (part of
the \model{AST} model calculation, Section~\ref{sec:ast}). In
Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
explicitly to zero beyond a radius of 60\,arcsec from the location of
Uranus (much larger than the FWHM of the main lobe), for all but the
final iteration. In this case, the map converges after 6 iterations,
and the ringing has been effectively removed. Since the map is now
flat away from the source, and constrained to a value of zero, it is
appropriate for performing aperture photometry directly, with no need
for an additional reference annulus. This approach to constrained
map-making is similar to that employed for poorly cross-linked scans
of compact (though resolved) sources by \citet{wiebe2009} using BLAST
data.
